Comparison of Fast Imaging Architectures for Large, Multi-scale, Hierarchical Aperture Arrays

### 2024-06-13 ms.tex
Update on Overleaf.

The same reasoning can be extended for inter-station processing, where there is significant diversity. There is a large range in the number of stations and in the array filling factor. As a result, both the layout and cadence interval play significant roles in determining the efficiency of architectures. For arrays with small numbers of stations such as \texttt{LAMBDA-I} and \texttt{CASPA}, the XBF architecture holds an advantage on all timescales. For arrays with large numbers of stations such as \texttt{SKA1-low-core} with a reasonably high packing density, EPIC holds the advantage on all timescales, but with slower cadence intervals, $t_\textrm{acc}\gtrsim 100$~ms, XBF and XFFT also have comparable efficiency. Although the \texttt{SKA1-low} has a large number of stations, they are not as densely packed as in the core, and therefore, the role of $t_\textrm{acc}$ is more significant. XFFT and EPIC are comparable at $t_\textrm{acc}\lesssim 1$~ms, but for $t_\textrm{acc}\gtrsim 100$~ms, XFFT is clearly advantageous. Similarly, \texttt{FarView} has a similar array filling factor as \texttt{SKA1-low-core}, but has far fewer stations. Hence, EPIC is only advantageous for $t_\textrm{acc}\lesssim 0.1$~ms, whereas XFFT becomes more efficient for $t_textrm{acc}\gtrsim 1$~ms.

@@

Besides computational costs, there are other practical considerations in the deployment of an architecture such as the hardware's processing bandwidth, internal memory bandwidth management, input and output data rates, etc.\footnote{\url{https://github.com/epic-astronomy/Memos/blob/master/PDFs/009_EPIC_Code_Optimizations.md}}

There are other computations involved that may be relevant for different applications such as deconvolution, dedispersion, etc. that may determine the exact architecture.

\cite{Sokolowski+2024}

Post-processing steps like deconvolution and de-dispersion add to these costs.

>>>>

The same reasoning can be extended for inter-station processing, where there is significant diversity. There is a large range in the number of stations and in the array filling factor. As a result, both the layout and cadence interval play significant roles in determining the efficiency of architectures. For arrays with small numbers of stations such as \texttt{LAMBDA-I} and \texttt{CASPA}, the XBF architecture holds an advantage on all timescales. For arrays with large numbers of stations such as \texttt{SKA1-low-core} with a reasonably high packing density, EPIC holds the advantage on all timescales, but with slower cadence intervals, $t_\textrm{acc}\gtrsim 100$~ms, XBF and XFFT also have comparable efficiency. Although the \texttt{SKA1-low} has a large number of stations, they are not as densely packed as in the core, and therefore, the role of $t_\textrm{acc}$ is more significant. XFFT and EPIC are comparable at $t_\textrm{acc}\lesssim 1$~ms, but for $t_\textrm{acc}\gtrsim 100$~ms, XFFT is clearly advantageous. Similarly, \texttt{FarView} has a similar array filling factor as \texttt{SKA1-low-core}, but has far fewer stations. Hence, EPIC is only advantageous for $t_\textrm{acc}\lesssim 0.1$~ms, whereas XFFT becomes more efficient for $t_\textrm{acc}\gtrsim 1$~ms.

@@

In transient search and cosmological applications, there will be additional processing required. For example, a de-dispersion trial process is required to discover new FRBs and pulsars with optimal signal-to-noise ratio. Computationally efficient algorithms for de-dispersion like the Fast Dispersion Measure Transform \citep[FDMT;][]{Zackay+2014} need to be performed on every independent pixel in the image formed. In intensity mapping applications for cosmology, the images need to be post-processed for deconvolution, foreground, removal, etc. As these post-processing computations are additional and common to all architectures, including them in our cost metric will not affect the results.

Besides computational costs, there are other practical considerations in the deployment of an architecture such as the hardware's processing bandwidth, internal memory bandwidth management, input and output data rates, etc. For example, the GPU-based deployment of EPIC on the LWA is severely limited by the memory bandwidth constraints on chip\footnote{See \url{https://github.com/epic-astronomy/Memos/blob/master/PDFs/009_EPIC_Code_Optimizations.md} for a detailed analysis.}. The architecture of current generation of GPUs is found to be inefficient at handling the transpose operation required in an $N$-point two-dimensional FFT, whereas they may be more efficient at handling correlations. On the contrary, the current generation of FPGAs seem more suited for low bit width two-dimensional FFT from computational and on-chip memory bandwidth perspectives. It is extremely difficult to forecast the continually evolving capabilities of hardware accessible to instruments at the instant of deployment. Thus, we restrict our analysis to the metric computational cost density, which is independent of evolving hardware capabilities. For example, even if the computations are parallelised over a large number of GPU cores as noted in \citet{Sokolowski+2024}, the total computational cost remains unaffected.
